---
title: "An Analysis of the English Premier League - Players and Teams 2016/2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
df <- read.csv("epldata_final.csv")
```

# Exploring the Dataset

This dataset contains information for 461 players of the English Premier League, participating in 2016/17. This is still incomplete, as the transfer window is open. The data was scraped from TransferMrkt.com and Fantasy Premier League predominantly. To see the scraping, click [here]( https://github.com/shubham-maurya/epl-data-scraping)

The head of the data looks like this - 

```{r}
head(df)
```


# What I'm trying to build

+ A regression model that explains a player's transfer value
+ Estimating the size of the EPL's impact on popularity of players who've been signed from abroad
+ Do a comparative analysis of the Top 6 vs other teams
+ Estimate the (inverse) English premium

But first, some nice and simple exploratory analysis.

# How are Market Values Distributed?

One would expect the distribution to be long-tailed on the right : Lots of cheap players, with much fewer quality players. 

```{r,warning=FALSE,echo=FALSE}
ggplot(df,aes(market_value))+geom_histogram(binwidth = 5)+xlab("Market Value (in Million Euros)")
```
Clearly, the majority of players are valued under 10 million euros.

## Top 6 (Yes, Tottenham too.)
```{r,warning=FALSE,echo=FALSE}
df %>% filter(big_club == 1) %>% 
ggplot(aes(market_value))+geom_histogram(binwidth = 5)+xlab("Market Value (in Million Euros)")
```

## Outside Top 6
```{r,warning=FALSE,echo=FALSE}
df %>% filter(big_club == 0) %>% 
ggplot(aes(market_value))+geom_histogram(binwidth = 5)+xlab("Market Value (in Million Euros)")
```

That gives us valuable information - the top 6 have a very different distribution as compared to the other clubs. (Okay, I'm going to clarify this right here - Top 6 is Manchester United, Tottenham, Manchester City, Arsenal, Chelsea, Liverpool) (in no particular order - okay fine. In the order I hope the 17/18 season finishes.)

# Distribution of Popularity
I've also scraped the popularity of individual players. While this is no easy task given the lack of objectivity of what 'popular' means, I decided to go with a classic. Number of Wikipedia searches. While I admit this is not perfect, I think it suffers less than Twitter and Facebook data - several players don't have an account, they aren't active on social media, and so on. A count of news mentions was interesting, but I couldn't find the appropriate data.

Wikipedia searches, in my mind, seems a passable choice. It is not prone to influence by newspapers and player's pushing their profile (see Pogba). Some players just have more interesting stories than others, but on the whole, I think it's safe to assume that Wikipedia reflects a player's popularity.

Now regarding the daterange for the searches - this was a bit of a headache. THe data archive only dated back to 2015. Anyway, I definitely wanted to avoid May - August because of the transfer speculation. I settled on using *average daily searches* from August 1, 2016 to May 1, 2017, giving the summer transfer window a miss. It would have been ideal to ignore the winter window as well, but there was no way to do so easily. Since it's a relatively quiet window, I decided to allow that (minimal) bias.

Now let's lok at the histograms of popularity - 

```{r,warning=FALSE,echo=FALSE}
ggplot(df,aes(page_views))+geom_histogram()+xlab("Average Daily Wiki Searches")
```

## Top 6 
```{r,warning=FALSE,echo=FALSE}
df %>% filter(big_club == 1) %>%filter(page_views < 6000 ) %>%
ggplot(aes(page_views))+geom_histogram()+xlab("Average Daily Wiki Searches")
```

## Outside Top 6
```{r,warning=FALSE,echo=FALSE}
df %>% filter(big_club == 0) %>% filter(page_views < 6000 ) %>%
ggplot(aes(page_views))+geom_histogram()+xlab("Average Daily Wiki Searches")
```

Herein lies my motivation for this project. There seems to be some correlation between market value and popularity, atleast at a highly-aggregated level like this. I think it's worth diving deeper into the data, to see if a decent model can be built. That'll be the focus of the work.